---
title: "IE 451 Homework 7"
author: "Mustafa Mert Türkmen"
format: 
  html:
    embed-resources: true
    toc: true
editor: visual
---
```{r}
#| echo: false
#| message: false
library(ISLR2)
library(leaps)
library(glmnet)
```
## 8


### 8.a

```{r}
set.seed(1907)
x <- rnorm(100)
ep <- rnorm(100)
```

### 8.b

```{r}
y <- 24 + 3*x + 1*x^2 + 5*x^3 + ep
```


### 8.c

```{r}
df <- data.frame(x,y)
reg <- regsubsets(y~poly(x,10,raw = TRUE, simple = TRUE), nvmax = 10, data = df)
summary(reg)
```
```{r}
cp_values <- summary(reg)$cp
bic_values <- summary(reg)$bic
adjr2_values <- summary(reg)$adjr2
```

```{r}
min_cp_index <- which.min(cp_values)
min_bic_index <- which.min(bic_values)
max_adjr2_index <- which.max(adjr2_values)
```


```{r}
par(mfrow = c(1, 3))
plot(cp_values, main = "Cp Values", type = "l", col = "blue", lwd = 2)
points(min_cp_index, cp_values[min_cp_index], col = "red", pch = 16)

plot(bic_values, main = "BIC Values", type = "l", col = "green", lwd = 2)
points(min_bic_index, bic_values[min_bic_index], col = "red", pch = 16)

plot(adjr2_values, main = "Adjusted R^2 Values", type = "l", col = "purple", lwd = 2)
points(max_adjr2_index, adjr2_values[max_adjr2_index], col = "red", pch = 16)
```
- With looking minimum Cp value and maximum AdjR2 values we can create a model with five variable.

```{r}
bic_values
```
- There is not much difference in bic_values at using 3 or 5 variable, so we can create best model with five variable according to these three parameters.

```{r}
coef(reg, id = 3)
```
- The coefficients as we would expect. 
- Intercept is close to 24.05 which is very close the 24.
- The model has contained X^1, X^2 and X^3, the values are close to our coefficients.

### 8.d

```{r}
reg_forward <- regsubsets(y~poly(x,10,raw = TRUE, simple = TRUE), nvmax = 10, data = df, method='forward')
summary(reg_forward)
```

```{r}
cp_values_f <- summary(reg_forward)$cp
bic_values_f <- summary(reg_forward)$bic
adjr2_values_f <- summary(reg_forward)$adjr2
```

```{r}
min_cp_index_f <- which.min(cp_values_f)
min_bic_index_f <- which.min(bic_values_f)
max_adjr2_index_f <- which.max(adjr2_values_f)
```


```{r}
par(mfrow = c(1, 3))
plot(cp_values_f, main = "Cp Values", type = "l", col = "blue", lwd = 2)
points(min_cp_index_f, cp_values_f[min_cp_index_f], col = "red", pch = 16)

plot(bic_values_f, main = "BIC Values", type = "l", col = "green", lwd = 2)
points(min_bic_index_f, bic_values_f[min_bic_index_f], col = "red", pch = 16)

plot(adjr2_values_f, main = "Adjusted R^2 Values", type = "l", col = "purple", lwd = 2)
points(max_adjr2_index_f, adjr2_values_f[max_adjr2_index_f], col = "red", pch = 16)
```

- The forward stepwise selection is changed our best subset approach, looking minimum Cp value and maximum AdjR2 values we can create a model with six variable. We add one more variable with this method.

```{r}
bic_values_f
```
- The BIC value is not lowest on the six variable however with looking cp and adjr2 we can choose six variable model as best subset with method forward.

```{r}
coef(reg_forward, id = 6)
```
- Intercept is close to 24.27 which is very close the 24.
- The model has contained X^1,X^2,X^3,X^4,X^6 and X^7.
- The coefficients are not close as much as first method but the coef of X^6 and X^7 is close to zero.
- I think BIC is better performed than other two evaluation method.

```{r}
reg_backward <- regsubsets(y~poly(x,10,raw = TRUE, simple = TRUE), nvmax = 10, data = df, method='backward')
summary(reg_backward)
```

```{r}
cp_values_b <- summary(reg_backward)$cp
bic_values_b <- summary(reg_backward)$bic
adjr2_values_b <- summary(reg_backward)$adjr2
```

```{r}
min_cp_index_b <- which.min(cp_values_b)
min_bic_index_b <- which.min(bic_values_b)
max_adjr2_index_b <- which.max(adjr2_values_b)
```


```{r}
par(mfrow = c(1, 3))
plot(cp_values_b, main = "Cp Values", type = "l", col = "blue", lwd = 2)
points(min_cp_index_b, cp_values_b[min_cp_index_b], col = "red", pch = 16)

plot(bic_values_b, main = "BIC Values", type = "l", col = "green", lwd = 2)
points(min_bic_index_b, bic_values_b[min_bic_index_b], col = "red", pch = 16)

plot(adjr2_values_b, main = "Adjusted R^2 Values", type = "l", col = "purple", lwd = 2)
points(max_adjr2_index_b, adjr2_values_b[max_adjr2_index_b], col = "red", pch = 16)
```
- The best subset suggested by backward selection is more consistent, because looking minimum Cp value, minimum BIC value and maximum AdjR2 values we can create a model with five variable. All of the values are on the same side right now. :)

```{r}
coef(reg_backward, id = 5)
```
- Intercept is close to 24.34 which is very close the 24.
- The model has contained X^1,X^3,X^4,X^6 and X^7.
- The coefficients are not close as much as first method also in the backward method.
- The X^2's coef is equal to zero, not similar with our model. 
- But the coef of X^6 and X^7 is close to zero.
- I think best performing method was the first one, which is default mode of regsubsets function which is 'exhaustive'.

### 8.e
```{r}
x_ten <- poly(x,10,raw = TRUE, simple = TRUE)
lasso_model <- cv.glmnet(y=y,x=x_ten, alpha = 1)
plot(lasso_model)
```
```{r}
optimal_lambda <- lasso_model$lambda.min
coefficients_lasso <- coef(lasso_model, s = optimal_lambda)
coefficients_lasso
```
- We can see that the coefficients for X4 - X10 have been shrunk to exactly zero.
- The coefficients for X - X3 are large, which makes sense as the generating function was an order 3 polynomial. 
- The intercept and coefficients for X, X^2 and X^3 closely match the ones chosen in 8(b).
- I chose 24, 3, 1, 5 respectively, lasso models results 25.17, 1.52, 0.26 and 5.06, which are very close values to mine.
- Therefore, we can say that this model provides an accurate estimation of the response Y.

### 8.f

```{r}
y1 <- 3 + 2*x^7 + ep
```

```{r}
df1 <- data.frame(x,y1)
reg1 <- regsubsets(y1~poly(x,10,raw = TRUE, simple = TRUE), nvmax = 10, data = df1)
summary(reg1)
```
```{r}
cp_values1 <- summary(reg1)$cp
bic_values1 <- summary(reg1)$bic
adjr2_values1 <- summary(reg1)$adjr2
```

```{r}
min_cp_index1 <- which.min(cp_values1)
min_bic_index1 <- which.min(bic_values1)
max_adjr2_index1 <- which.max(adjr2_values1)
```


```{r}
par(mfrow = c(1, 3))
plot(cp_values1, main = "Cp Values", type = "l", col = "blue", lwd = 2)
points(min_cp_index1, cp_values1[min_cp_index1], col = "red", pch = 16)

plot(bic_values1, main = "BIC Values", type = "l", col = "green", lwd = 2)
points(min_bic_index1, bic_values1[min_bic_index1], col = "red", pch = 16)

plot(adjr2_values1, main = "Adjusted R^2 Values", type = "l", col = "purple", lwd = 2)
points(max_adjr2_index1, adjr2_values1[max_adjr2_index1], col = "red", pch = 16)
```
- With looking minimum Cp value and minimum BIC values we can create a model with one variable.
- The matrix from summary of regsubsets showing include X7 for single variable model. (same as our generated function)


```{r}
coef(reg1, id = 1)
```
- The coefficient as we would expect. 
- Intercept is close to 3.08 which is very close the 3.
- The model has contained X^7 and the value is so close to our coefficient which is 2.

```{r}
x_ten1 <- poly(x,10,raw = TRUE, simple = TRUE)
lasso_model1 <- cv.glmnet(y=y1,x=x_ten1, alpha = 1)
plot(lasso_model1)
```
```{r}
optimal_lambda1 <- lasso_model1$lambda.min
coefficients_lasso1 <- coef(lasso_model1, s = optimal_lambda1)
coefficients_lasso1
```
- The coefficent as expected but the intercept point is not. 
- Intercept is much higher than 3.
- The model has contained X^7 and the value is so close to our coefficient which is 2.

## 9
### 9.a


```{r}
set.seed(1907)
train_indices <- sample(1:nrow(College), nrow(College) * 0.75)
train_set <- College[train_indices, ]
test_set <- College[-train_indices, ]
```

### 9.b
- The lm function used to perform least squares linear regression so I can use that.

```{r}
lm <- lm(Apps~., train_set)
summary(lm)
```

```{r}
pred <- predict(lm, newdata = test_set)
test_error <- mean((pred - test_set$Apps)^2)
test_error
```
### 9.c
```{r}
x_train <- model.matrix(Apps ~ ., data = train_set)
y_train <- train_set$Apps
x_test <- model.matrix(Apps ~ ., data = test_set)
```


```{r}
ridge_model <- cv.glmnet(x_train,y_train, alpha = 0)
ridge_model
```

```{r}
lambda <- ridge_model$lambda.min
lambda
```

```{r}
ridge <- glmnet(x_train, y_train, alpha = 0, lambda = lambda)
ridge_pred <- predict(ridge, s = lambda, newx = x_test)
ridge_mse <- mean((ridge_pred - test_set$Apps)^2)
ridge_mse
```
- Best value of lambda is 310.7218 and the test MSE is higher than for least squares.

### 9.d
- We chose alpha equals 1 to create lasso regression
```{r}
lasso_model <- cv.glmnet(x_train, y_train, alpha = 1) 
```

```{r}
lasso_lambda <- lasso_model$lambda.min
lasso_lambda
```
```{r}
lasso <- glmnet(x_train, y_train, alpha = 1, lambda = lasso_lambda)
lasso_pred <- predict(lasso, s=lasso_lambda, newx = x_test)
lasso_mse <- mean((lasso_pred - test_set$Apps)^2)
lasso_mse
```
- Best value of lambda is 11.69848 and the test MSE is higher than for least squares lower than ridge.

```{r}
nonzero_lasso <- sum(coef(lasso, s = lasso_lambda) != 0)
nonzero_lasso
```
```{r}
coef(lasso, s = lasso_lambda)
```
- The total number of the non-zero coefficients is 14 out of 19.
