---
title: "IE 451 Homework 11"
author: "Mustafa Mert Türkmen"
format: 
  html:
    embed-resources: true
    toc: true
editor: visual
---
```{r}
#| echo: false
#| message: false
#| warning: false
library(ISLR2)
library(MASS)
library(e1071)
library(class)
library(randomForest)
library(magrittr)
library(glmnet)
library(tidyverse)
```

## 14

### 14.a
```{r}
x <- cbind(Auto[, -1], data.frame("mpg01" = Auto$mpg > median(Auto$mpg)))
```


### 14.b
```{r}
par(mfrow = c(2, 4))
for (i in 1:7) boxplot(x[, i] ~ x$mpg01, main = colnames(x)[i])
```
- Most variables show a relation with mpg01 category. 
- If mpg is above the median;
  -   Cylinders generally 4
  -   Displacement is lower than below median
  -   Horsepower is lower than below median, kinda surprising
  -   Horsepower is higher
  -   They are generally new
  
### 14.c
```{r}
set.seed(42)
train <- sample(seq_len(nrow(x)), nrow(x) * 0.7)
```

### 14.f

```{r}
logit_model  <- glm(mpg01 ~ cylinders + displacement + horsepower + weight + acceleration, data = x[train, ], family = binomial)
logit_predictions  <- predict(logit_model, x[-train, ], type = "response") > 0.5
logit_error <- mean(logit_predictions != x[-train, ]$mpg01)
logit_error
```
```{r}
#| fig.asp: 1
roc_df_logit <- x[train, ] %>% 
  select(mpg01) %>%
  mutate(prob = predict(logit_model, type = "response" )) %>%
  arrange(desc(prob)) %>% 
  mutate(TP = cumsum(ifelse(mpg01 == 1, 1, 0)),
         FP = cumsum(ifelse(mpg01 == 0, 1, 0)),
         TPR = TP / sum(mpg01 == 1),
         FPR = FP / sum(mpg01 == 0))

roc_df_logit %>% 
  ggplot(aes(FPR, TPR)) +
  geom_point() +
  geom_hline(yintercept = c(0, 1), lty = "dashed") +
  geom_vline(xintercept = c(0, 1), lty = "dashed") +
  geom_abline(lty = "dashed") +
  labs(title = "ROC Curve",
       x = "False Positive Rate",
       y = "True Positive Rate")
```
```{r}
#| warning: false
roc_df_logit %>% summary
roc_logit <- approxfun(roc_df_logit$FPR, roc_df_logit$TPR)
curve(roc_logit)
roc_logit(0.001)
integrate(roc_logit, 0.0001, 1)$value
```
- The AUC for logistic regression is 0.951224

### Random Forest

```{r}
rf_model <- randomForest(factor(mpg01) ~ cylinders + displacement + horsepower + weight + acceleration + year,
                          data = x[train, ])
rf_pred <- predict(rf_model, newdata =  x[-train, ])
rf_error <- mean(rf_pred != x[-train, ]$mpg01)
rf_error
```
```{r}
#| fig.asp: 1
roc_df_rf <- x[train, ] %>% 
  select(mpg01) %>%
  mutate(prob = predict(rf_model, type = "response" )) %>%
  arrange(desc(prob)) %>% 
  mutate(TP = cumsum(ifelse(mpg01 == 1, 1, 0)),
         FP = cumsum(ifelse(mpg01 == 0, 1, 0)),
         TPR = TP / sum(mpg01 == 1),
         FPR = FP / sum(mpg01 == 0))

roc_df_rf %>% 
  ggplot(aes(FPR, TPR)) +
  geom_point() +
  geom_hline(yintercept = c(0, 1), lty = "dashed") +
  geom_vline(xintercept = c(0, 1), lty = "dashed") +
  geom_abline(lty = "dashed") +
  labs(title = "ROC Curve",
       x = "False Positive Rate",
       y = "True Positive Rate")
```
```{r}
#| warning: false
roc_df_rf %>% summary
roc_rf <- approxfun(roc_df_rf$FPR, roc_df_rf$TPR)
curve(roc_rf)
roc_rf(0.001)
integrate(roc_logit, 0.0001, 1)$value
integrate(roc_rf, 0.0001, 1)$value
```
- Random forest AUC = 0.96295 which is higher than logit AUC
- Based on AUC, Random forest is better than logistic regression.

### Lasso Logistic Regression

```{r}
lasso.mod <- cv.glmnet(as.matrix(x[train,][,-c(8,9)]),x[train, ]$mpg01,alpha=1)
lasso_pred <- predict(lasso.mod, s = "lambda.min", newx = as.matrix(x[-train, ][, -c(8,9)]))
lasso_pred_binary <- ifelse(lasso_pred > 0.5, 1, 0)
lasso_error <- mean(lasso_pred_binary != x[-train, ]$mpg01)
lasso_error
```
```{r}
#| fig.asp: 1
roc_df_lasso <- x[train, ] %>%
  select(mpg01) %>%
  mutate(prob = predict(lasso.mod, newx = as.matrix(x[train, ][, -c(8, 9)]), type = "response")) %>%
  arrange(desc(prob)) %>%
  mutate(TP = cumsum(ifelse(mpg01 == 1, 1, 0)),
         FP = cumsum(ifelse(mpg01 == 0, 1, 0)),
         TPR = TP / sum(mpg01 == 1),
         FPR = FP / sum(mpg01 == 0))

roc_df_lasso %>%
  ggplot(aes(FPR, TPR)) +
  geom_point() +
  geom_hline(yintercept = c(0, 1), lty = "dashed") +
  geom_vline(xintercept = c(0, 1), lty = "dashed") +
  geom_abline(lty = "dashed") +
  labs(title = "ROC Curve",
       x = "False Positive Rate",
       y = "True Positive Rate")
```

```{r}
#| warning: false
roc_df_lasso %>% summary
roc_lasso <- approxfun(roc_df_lasso$FPR, roc_df_lasso$TPR)
curve(roc_lasso)
roc_lasso(0.001)
integrate(roc_logit, 0.0001, 1)$value
integrate(roc_rf, 0.0001, 1)$value
integrate(roc_lasso, 0.0001, 1)$value
```
- Lasso reg AUC = 0.960545 which is higher than logit AUC
- Based on AUC, Random forest is better than lasso regression.


### Ridge Logistic Regression

```{r}
ridge.mod <- cv.glmnet(as.matrix(x[train,][,-c(8,9)]),x[train, ]$mpg01,alpha=0)
ridge_pred <- predict(ridge.mod, s = "lambda.min", newx = as.matrix(x[-train, ][, -c(8,9)]))
ridge_pred_binary <- ifelse(ridge_pred > 0.5, 1, 0)
ridge_error <- mean(ridge_pred_binary != x[-train, ]$mpg01)
ridge_error
```
```{r}
#| fig.asp: 1
roc_df_ridge <- x[train, ] %>%
  select(mpg01) %>%
  mutate(prob = predict(ridge.mod, newx = as.matrix(x[train, ][, -c(8, 9)]), type = "response")) %>%
  arrange(desc(prob)) %>%
  mutate(TP = cumsum(ifelse(mpg01 == 1, 1, 0)),
         FP = cumsum(ifelse(mpg01 == 0, 1, 0)),
         TPR = TP / sum(mpg01 == 1),
         FPR = FP / sum(mpg01 == 0))

roc_df_ridge %>%
  ggplot(aes(FPR, TPR)) +
  geom_point() +
  geom_hline(yintercept = c(0, 1), lty = "dashed") +
  geom_vline(xintercept = c(0, 1), lty = "dashed") +
  geom_abline(lty = "dashed") +
  labs(title = "ROC Curve",
       x = "False Positive Rate",
       y = "True Positive Rate")
```

```{r}
#| warning: false
roc_df_ridge %>% summary
roc_ridge <- approxfun(roc_df_ridge$FPR, roc_df_ridge$TPR)
curve(roc_ridge)
roc_ridge(0.001)
integrate(roc_logit, 0.0001, 1)$value
integrate(roc_rf, 0.0001, 1)$value
integrate(roc_lasso, 0.0001, 1)$value
integrate(roc_ridge, 0.0001, 1)$value
```
- Lasso reg AUC = 0.9564662 which is higher than logit AUC
- Based on AUC, Random forest is better than ridge regression also.

- I would choose the Random Forest model based on its superior performance in terms of both AUC scores and ROC curves. The Random Forest model consistently demonstrated higher AUC values, indicating better overall predictive ability compared to the other models.